---
title: "TEOS Lab Using Hydrostation S Data"
author: "Carlos J. Arroyo Ruiz"
date: "2023-03-07"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Load required libraries
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(gsw)
library(plotly)

```

## Now we need to import our data
```{r warning=FALSE, message=FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)

hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)

colnames(hydrostation_bottle)=colnames(hydrostation_bottle_names)

#View(hydrostation_bottle)
```
## Variable names and units
- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
- Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3)       
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    
- /Quality flags
- -999  = No data
- 0 = Less than detection limit
 
```{r}
#Let's first plot the data.
hydrostation_bottle %>%
  ggplot()+geom_point(aes(x=decy,y=`Sig-th`))

#This was illogical, so filter out the -999 no data flag.
hydrostation_bottle %>%
  filter(`Sig-th`!=-999) %>%
  ggplot()+geom_point(aes(x=decy,y=`Sig-th`))

#This looks much better but still hard to interpret, so let's only look at the surface data.
hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+geom_point(aes(x=decy,y=`Sig-th`))

#Still not super clear, so lets try a line plot.
hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+geom_line(aes(x=decy,y=`Sig-th`))

#Clear seasonal signal for sigma-theta, lets see how this compares to temperature.
hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+geom_point(aes(x=Temp,y=`Sig-th`))

#Temperature and density are strongly correlated, but there appears to be two outliars that we ill likely need to address at some point.
#We only have density data from 1988-present, but temperature and salinity data from the 1950's-present.
#We can use the TEOS-10 to do this.
```
 
## TEOS-10 Toolbox in Package seacarb

```{r}
?gsw #Launches the documentation for the Gibbs seawater toolbox (TEOS-10)
?gsw_sigma0 #Let's check this function. It says we need absolute salinity and conservative temperature.

#First we need absolute salinity:
?gsw_SA_from_SP
#Practical salinity
#Sea pressure (dbar)
#Longitude
#Latitude

#Let's plot our pressure data - it's missing before 1980's.
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy,y=Pres))

#We have depth data for the time series.
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy,y=Depth))

?gsw_p_from_z

#Adds a pressure column from the depth and latN columns from/to hydrostation bottle
hydrostation_bottle=
  hydrostation_bottle %>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN))

hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=Pres,y=Pres_gsw))
#We see strong 1:1 agreement between measured pressure and calculated pressure.

#Checking lat, lon, and salinity data.
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy,y=Sal1))

hydrostation_bottle=
  hydrostation_bottle %>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN))%>%
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))

#Check it!
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy,y=S_abs_gsw))

#How else can I check my data?
hydrostation_bottle %>%
  filter(Sal1!=-999)%>%
  ggplot()+
  geom_point(aes(x=Sal1,y=S_abs_gsw))

#Now we need to calculate conservative temperature.
?gsw_CT_from_t
#We need absolute salinity, in-situ temp(ITS-90), and sea pressure.

#Add line to calculate conservative temperature.
hydrostation_bottle=
  hydrostation_bottle %>%
  filter(Sal1!=-999)%>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN))%>%
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))%>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))

#Let's check our data.
hydrostation_bottle %>%
  filter(Temp!=-999)%>%
  ggplot()+
  geom_point(aes(x=Temp,y=T_cons_gsw))

HydroS=
  hydrostation_bottle %>%
  filter(Sal1!=-999)%>%
  filter(Temp!=-999) %>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN))%>%
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))%>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))

HydroS=
  hydrostation_bottle %>%
  filter(Sal1!=-999)%>%
  filter(Temp!=-999) %>%
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1,latN))%>%
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))%>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))%>%
  mutate(Sig_th_gsw=gsw_sigma0(S_abs_gsw,T_cons_gsw))

HydroS %>%
  filter(`Sig-th`!=-999)%>%
  ggplot()+
  geom_point(aes(x=`Sig-th`,y=Sig_th_gsw))
#But we have a very low Sig-th-gsw so let's find it.

HydroS %>%
  filter(Sig_th_gsw<0)%>%
  View()

HydroS_correctedS_a=
  HydroS%>%
  filter(Sig_th_gsw<0)%>%
  mutate(S_abs_gsw=gsw_SA_from_SP(CTD_S,Pres_gsw,360-lonW,latN))%>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))%>%
  mutate(Sig_th_gsw=gsw_sigma0(S_abs_gsw,T_cons_gsw))

#add spiciness mutate line
HydroS_correctedS_b=
  HydroS%>%
  filter(Sig_th_gsw>0)

HydroS_corrected=rbind(HydroS_correctedS_a,HydroS_correctedS_b)

HydroS_corrected %>%
  filter(`Sig-th`!=-999)%>%
  ggplot()+
  geom_point(aes(x=`Sig-th`,y=Sig_th_gsw))

#Homework: Unite Hydro_corretedS_a and Hydro_corretedS_b script into original HydroS script.

HydroS_corrected %>%
  ggplot()+
  geom_point(aes(x=Sig_th_gsw,y=Depth))+
  scale_y_reverse()+
  scale_x_continuous(position="top")+
  xlab(expression(paste(sigma[theta],"(kg m"^"-3",")")))+
  ylab("Depth (m)")+
  theme_classic()
```

# Has surface sigma theta decreased over time?

```{r}
HydroS_shallow=HydroS_corrected %>%
  filter(Depth<30)

?lm
#lm(y~x, data=data)
lm(Sig_th_gsw~decy,data=HydroS_shallow)
#Coefficients (intercept and decy)
#y = mx+b
#y = Sig_th_gsw
#x = decy
#Coeficcients: intercept = b, decy = m
#Sig_th_gsw = -0.004*decy + 33.4
#(kg/m^3) = (kg/m^3/y)*y + (kg/m^3)
lm(Sig_th_gsw~decy,data=HydroS_shallow)
Sig_theta_time_model=lm(Sig_th_gsw~decy,data=HydroS_shallow)
summary(Sig_theta_time_model)

#library(plotly)
plot=HydroS_shallow %>%
  ggplot(aes(x=decy,y=Sig_th_gsw))+
  geom_point()+
  geom_line()+
  geom_smooth(method="lm")+
  theme_classic()
  
ggplotly(plot)
```

#Lab Assigment
1. Pick a question (include hypothesis).
2. Produce a plot and a statistical summary using lm()
3. Describe your results, the summary, and answer the question.
4. Compile into a completed lab report using R Markdown.

Potential questions:
  How do temperature, salinity, and salinity and sigma-theta co-vary?
  Is there a relationship between sigma-theta and oxygen?
  Is there a relationship of any of the parameters with depth? With time? Within a depth range over time?
  Are there seasonal differences in any of the parameters?
  
```{r}
#In Bermuda:
#Dec to March = Winter months
#August to October = summer months
HydroS_seasons=
  HydroS_corrected %>%
  mutate(month=as.numeric(substr(yyyymmd,5,6))) %>%
  mutate(season=ifelse(month==12|month==1|month==2|month==3,'winter',
         ifelse(month==8|month==9|month==10,'summer',NA)))

#Is dissolved oxygen higher in summer than winter(in shallow waters)?
#Is salinity higher in summer than winter?
#Is sound speed higher...
#Is spiciness higher...

summary(lm(`O2(1)`~season,data=HydroS_seasons))


#HydroS_months=
  #HydroS_corrected %>%
  #mutate(month=as.numeric(substr(yyyymmd,5,6)))
#View(HydroS_months)

#month>=8 & month<=10
#HydroS_months$season=
  #months==c(12,1,2,3)


```
  
  